Einstein's gravity from a polynomial affine model

### 2016-05-09 AffGR-PRD.tex
Corrected abstract, part of the introduction

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds---with or without cosmological constant---as solutions. Moreover, the effective field equations coincide with some of those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model. Finally, we present a naive analysis of the propagators of the gravitational theory, an their relation with the presence of ghost in the model. For completeness we include three appendices, including one in which it is shown that a Birkhoff-like theorem is valid for the analyzed sector.
\end{abstract}

@@

A hundred years ago, A.~Einstein presented by the first time the correct field equations of its relativistic theory of gravitation, known as General Relativity~\cite{einstein1915feldgleichungen}. It was not only consistent with the available observations, but it also predicted several effects that have been consistently observed and measured---see Ref.~\cite{Will:2014kxa} for a recent review---, being the last one the confirmation of gravitational waves~\cite{Abbott:2016blz}. Thus, General Relativity has become the most successful theory of gravitational interactions nowadays.

However, General Relativity seems to be an effective theory of a, yet to be discovered, more fundamental one. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity. In addition, we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization, General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Interestingly, gravitational theories with generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe}, or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}. The last sentence, allows to argue that a model cannot be discarded  for the presence of ghosts model, without a rigorous analysis of its stability.

@@

\begin{widetext}
  The perturbed action is then
  \begin{dmath}[compact, spread=2pt]
  \label{pert-Lambda}
    S = \int \dn{4}{x} \Bigg(
    C_1 \Big( {t_{\mathtt{s}}}^{\alpha\beta }-\frac{1}{4}g^{\alpha\beta}{{t_{\mathtt{s}}}^{\gamma}{}_{\gamma}} \Big) \bigg( -  (\frac{4\Lambda^2}{3}+\Lambda \Delta+\Delta^2) {\gamma_{\mathtt{s}}}_{\alpha\beta}  
    - \frac{8}{3}   \Lambda\Delta {{\gamma}_{\alpha\beta}} 
    \bigg)
    + {{t_{\mathtt{v}}}^{\alpha}}\bigg(
    \Big(( C_1 - 2\, C_2) \Delta^3 + \frac{1}{3}( C_1 + 2C_2 )\Lambda\Delta^2 + \frac{2}{3}( 2C_1 +  C_2 )\Lambda^2\Delta - \frac{1}{3}C_1  \Lambda^3\Big){{\gamma_{\mathtt{v}}}_{\alpha}}
    + \Big((C_1 - 2\, C_2) \Lambda\Delta - 2\, C_2 \Delta^2\Big){{\gamma_\perp}^{\beta}\,_{\beta \alpha}}
    + 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} \Delta{\nabla}^{\beta }\Delta{{t_{\mathtt{a}}}^{\gamma\delta}}  
    +\Big(- 2\, E_2 \Delta^2 +  (- E_2 + B_3 - \frac{2}{3}\, B_4) \Lambda\Delta\Big){a_\perp}_{\alpha} 
    \bigg)
    + {{t_{\mathtt{s}}}^{\alpha}{}_{\alpha}}\bigg( \frac{2}{3}C_1 \Lambda\Delta{{\gamma}^{\beta}\,_{\beta}} 
    + \frac{2}{3}E_2 \Lambda \Delta{a_{\mathtt{L}}}  
    - \frac{4}{3}C_1 \Lambda^2 \Delta{g_{\mathtt{L}}}
    - \frac{1}{4}C_1 (\Lambda + \Delta)  \Delta{{\gamma_{\mathtt{s}}}^{\beta}{}_{\beta}}
    \bigg)   \Bigg),
  \end{dmath}
  \hl{1.- Is it Ok?}
\end{widetext}
where $\Delta \equiv \nabla^{\alpha}\nabla_{\alpha}$ is the Laplacian operator. This action sets the dynamics of two transverse-traceless tensors (or specific combination of tensors), as well as two transverse vectors and two scalars, which are associated to eighteen degrees of freedom.

@@

%%%%%%%%% ACKNOWLEGMENTS %%%%%%%%%
\begin{acknowledgments}
  We thank to N.~Pantoja, A.~Melfo, M.~Blagojevi\'c, B.~Cvetkovi\'c, V.~Sharma, O.~Orellana and R.~L.~Bryant for their helpful discussions and inspiring comments, to J.~Zanelli for his suggestions on the physical insight into the problem and careful but critical review of the manuscript.
  This work was partially supported by CONICYT (Chile) under project No. 79140040.
\end{acknowledgments}

>>>>

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds---with or without cosmological constant---as solutions. Moreover, the effective field equations are partially those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity within a ``minimally'' coupled scalar field in this affine model. Finally, we present a brief analysis of the propagators of the gravitational theory, and count the degrees of freedom. For completeness we include three appendices, including one in which we show that a Birkhoff-like theorem is valid for the analyzed sector.
\end{abstract}

@@

A hundred years ago, A.~Einstein presented by the first time the correct field equations of its relativistic theory of gravitation, known as General Relativity~\cite{einstein1915feldgleichungen}. It was not only consistent with the available observations, but it also predicted several effects that have been consistently observed and measured---see Ref.~\cite{Will:2014kxa} for a recent review---, being the last one the confirmation of the existence of the existence of gravitational waves~\cite{Abbott:2016blz}. Thus, General Relativity has become the most successful theory of gravitational interactions nowadays.

However, General Relativity seems to be an effective theory of a, yet to be discovered, more fundamental one. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity (when it is coupled to matter). In addition, we mention a few aspects related with the quantization: First, the fact that within the standard formalism of quantization, General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. Finally, it is well known that once matter couples to gravity (in particular fermionic matter), the assumption of vanishing torsion in standard General Relativity is no longer preserved as a field equation, and the best approach is to consider Cartan's generalization of General Relativity~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.
%%it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Interestingly, gravitational theories with a generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe}, or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}. The last sentence, allows to argue that a model cannot be discarded  for the presence of ghosts model, without a rigorous analysis of its stability.

@@

\begin{widetext}
  The perturbed action is then
  \begin{dmath}[compact, spread=2pt]
  \label{pert-Lambda}
    S = \int \dn{4}{x} \Bigg(
    C_1 \Big( {t_{\mathtt{s}}}^{\alpha\beta }-\frac{1}{4}g^{\alpha\beta}{{t_{\mathtt{s}}}^{\gamma}{}_{\gamma}} \Big) \bigg( -  (\frac{4\Lambda^2}{3}+\Lambda \Delta+\Delta^2) {\gamma_{\mathtt{s}}}_{\alpha\beta}  
    - \frac{8}{3}   \Lambda\Delta {{\gamma}_{\alpha\beta}} 
    \bigg)
    + {{t_{\mathtt{v}}}^{\alpha}}\bigg(
    \Big(( C_1 - 2\, C_2) \Delta^3 + \frac{1}{3}( C_1 + 2C_2 )\Lambda\Delta^2 + \frac{2}{3}( 2C_1 +  C_2 )\Lambda^2\Delta - \frac{1}{3}C_1  \Lambda^3\Big){{\gamma_{\mathtt{v}}}_{\alpha}}
    + \Big((C_1 - 2\, C_2) \Lambda\Delta - 2\, C_2 \Delta^2\Big){{\gamma_\perp}^{\beta}\,_{\beta \alpha}}
    + 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} \Delta{\nabla}^{\beta }\Delta{{t_{\mathtt{a}}}^{\gamma\delta}}  
    +\Big(- 2\, E_2 \Delta^2 +  (- E_2 + B_3 - \frac{2}{3}\, B_4) \Lambda\Delta\Big){a_\perp}_{\alpha} 
    \bigg)
    + {{t_{\mathtt{s}}}^{\alpha}{}_{\alpha}}\bigg( \frac{2}{3}C_1 \Lambda\Delta{{\gamma}^{\beta}\,_{\beta}} 
    + \frac{2}{3}E_2 \Lambda \Delta{a_{\mathtt{L}}}  
    - \frac{4}{3}C_1 \Lambda^2 \Delta{g_{\mathtt{L}}}
    - \frac{1}{4}C_1 (\Lambda + \Delta)  \Delta{{\gamma_{\mathtt{s}}}^{\beta}{}_{\beta}}
    \bigg)   \Bigg),
  \end{dmath}
  %% \hl{1.- Is it Ok?}
\end{widetext}
where $\Delta \equiv \nabla^{\alpha}\nabla_{\alpha}$ is the Laplacian operator. This action sets the dynamics of two transverse-traceless tensors (or specific combination of tensors), as well as two transverse vectors and two scalars, which are associated to eighteen degrees of freedom.

@@

%%%%%%%%% ACKNOWLEGMENTS %%%%%%%%%
\begin{acknowledgments}
  We thank to N.~Pantoja, A.~Melfo, M.~Blagojevi\'c, B.~Cvetkovi\'c, C.~Corral, V.~Sharma, O.~Orellana and R.~L.~Bryant for their helpful discussions and inspiring comments, to J.~Zanelli for his suggestions on the physical insight into the problem and careful but critical review of the manuscript.
  This work was partially supported by CONICYT (Chile) under project No. 79140040.
\end{acknowledgments}
